---
title: "Assignment_B4"
author: "Jérôme Plumier"
date: "2023-12-07"
output: github_document
---

# Option A - Strings and functional programming in R

For the completion of the fourth assignment for STAT545B, option A was chosen.
To complete option A, exercise 1 and exercise 2 will be done. 

First, let's load some libraries we will use.

```{r}
library(tidyverse)
library(dplyr)
library(tidytext)
library(ggplot2)
library(stringr)
library(roxygen2)
library(testthat)
```

## Exercise 1 - Most common words (37.5 points)

For the completion of exercise 1, we shall use the Jane Austen books, using the
`janeaustenr` and `stopwords`package.

```{r}
library(janeaustenr)
library(stopwords)
```

Now, we create our new dataframe `books_words` that contains Jane Austen's
6 books.

```{r}
books<-austen_books() # Place all 6 books in a books dataframe

# Make a new dataframe where each observation is a single word
books_words <- unnest_tokens(books,            # Select our Jane Austen book dataframe
                             word,            # Name the output as words
                             text,             # Input is the variable named text
                             token = "words",  # Name the output words
                             to_lower = TRUE,  # Make all words lower case
                             drop = TRUE)      # Drop the original text variable
head(books_words)
```

This permits us to create another dataframe named `stopwords_austen` with all
the English stopwords provided by the `stopwords`package removed. 
We can see, comparing with the previous `head()` function results, 
that the stopwords were removed.

```{r}
stopwords_austen <- books_words %>%
  filter(!word %in% stopwords("en"))
head(stopwords_austen)
```

To obtain our finalised dataframe, we now want to add the count of each words.

```{r}
count_austen <- stopwords_austen %>%
  group_by(book, word) %>%
  summarise(.groups="keep", n=n()) 
tail(count_austen)
```

We can now plot the most common words. For the sake of legibility, we will plot 
only the ten most common words. We will use as an example all six books first:

```{r}
plot_austen <- ggplot(data = (count_austen %>%
  arrange(desc(n)) %>% head(n=10))) +
  geom_bar(aes(y=word, x=n, fill = word), stat= "identity") +
  theme_bw() +
  theme(legend.position="none") +
  ylab("Ten most common words") +
  xlab("Number of word appearances")+
  ggtitle("All six Jane Austen books")
plot_austen
```

And as another example, here we looked at the ten most common words in the 
`Sense & Sensibility` book:

```{r}
plot_austen_SandS <- ggplot(data = (count_austen %>%
  filter(book =="Sense & Sensibility") %>%
  arrange(desc(n)) %>% head(n=10))) +
  geom_bar(aes(y=word, x=n, fill = word), stat= "identity") +
  theme_bw() +
  theme(legend.position="none") +
  ylab("Ten most common words") +
  xlab("Number of word appearances")+
  ggtitle("Sense & Sensibility")
plot_austen_SandS
```

## Exercise 2 - Pig Latin function (37.5 points)

This section will be further divided into three subsections.

### Exercise 2.1: Make a function

Let's first create a function that will convert words into my own Pig Latin 
version, and document it using `roxygen2`. 

It has a rearrangement component:
For words that end with a vowel, that vowel is duplicated at the beginning of 
the word sequence. In all other cases, the end of a word sequence from its 
last vowel onward is placed at the beginning of the word sequence.

It has an addition component:
All words receive a "o" that follows the new added beginning of the word 
sequence, and a "y" that comes at the new end of the word sequence.

```{R}

#' @title Transform a singular word into its pig Latin equivalent
#' 
#' @description Given a string of letters, this function will return a pig Latin
#' word sequence based on personal rules: For words that end with a vowel, that
#' vowel is duplicated at the beginning of the word sequence. In all other 
#' cases, the end of a word sequence from its last vowel onward is placed at
#' the beginning of the word sequence. Moreover, all words receive a "o" that
#' follows the new added beginning of the word sequence, and a "y" that comes
#' at the end of the new word sequence.
#' 
#' @param input, a string. It must be composed solely out of letters, and
#' contain both at least one vowel and one consonant.
#' 
#' @return a new word sequence with only lower case letters. It is the 
#' inputed string modified by personal pig Latin rules.
#' 
#' @export

pigLatin <- function(input) {
  
  if(!(is.character(input))) {
    stop('The word sequence needs to be inputed as a string.')}
  if(!grepl("^[A-Za-z]+$", input)) {
    stop('The word sequence needs to be made out of letters only (no space).')}

  input<-tolower(input)
  vowel_position <- grep("[aeiouy]", strsplit (input, NULL)[[1]])

  if(length(vowel_position)==0){
    stop('The word sequence contains no vowel.')}
  consonant_check <- grep("[zrtpqsdfghjklmwxcvbn]", strsplit (input, NULL)[[1]])
  if(length(consonant_check)==0){
    stop('The word sequence contains only vowels.')}

  vowel_position <- tail(vowel_position, 1)
  if(str_sub(input, start=-1)=="vowel_position"){
    str_c(str_sub(input, start = vowel_position), "o", str_sub(input, start = 1L, end = 1L), "y") 
  }else{
    str_c(str_sub(input, start = vowel_position), "o", str_sub(input, start = 1L, end = vowel_position-1), "y")
  }
}
```

### Exercise 2.2: Function examples

Here is an example of using the function to transform the word`hello`, which
ends with a vowel:

```{R}
test1 = "hello"
pigLatin(test1)
```

Here is an example of using the function to transform the word`transform`, which
ends with a consonant:

```{R}
test2 = "transform"
pigLatin(test2)
```

Here is an example of using the function to transform the word`HAMBURGER`, with 
only capital letters:

```{R}
test3 = "HAMBURGER"
pigLatin(test3)
```

### Exercise 2.3: Function tests

For this exercise, I tested my function using the `expect_error()` function 
from the `testthat` package with non-redundant inputs. They all passed:

```{R}
test_that('Not a string', {
x=c("a", "b", "c")
expect_error(pigLatin(x))
})
```

```{R}
test_that('Only one word', {
x="Hello world"
expect_error(pigLatin(x))
})
```

```{R}
test_that('The string is only made out of letters', {
x="Hello1"
expect_error(pigLatin(x))
})
```

```{R}
test_that('The string has at least one consonant', {
x="aeouiy"
expect_error(pigLatin(x))
})
```

```{R}
test_that('The string has at least one vowel', {
x="zrtpqsdfg"
expect_error(pigLatin(x))
})
```
